---
title: "06_make_phenofiles"
author: "Heather E. Wheeler"
date: "`r Sys.time()`"
output: html_document
---

###Load packages, functions, paths
```{r load,message=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(sas7bdat) ##read SAS files
library(foreign) ##read SPSS files, doesn't work with Fossa file, ask Ryan to convert to txt
library(psych)
library(MASS)
"%&%" = function(a,b) paste(a,b,sep="")
elementMean <- function(my.list) { 
  arr <- array(unlist(my.list),c(dim(my.list[[1]])[1],dim(my.list[[1]])[2],length(my.list)))
  rowMeans( arr , na.rm=TRUE, dims = 2 )
}
source("/Volumes/dolan-lab/hwheeler/ThePlatinumStudy/CIPN_modeling/ztransform.R")
source("/Volumes/dolan-lab/hwheeler/ThePlatinumStudy/CIPN_modeling/rntransform.R")
my.dir = "/Volumes/dolan-lab/hwheeler/ThePlatinumStudy/GWAS/"
pca.dir = "/Volumes/dolan-lab/hwheeler/ThePlatinumStudy/GWAS/PCA/"
pheno.dir = "/Volumes/dolan-lab/hwheeler/ThePlatinumStudy/GWAS/Datasets_from_Ryan_Cook_20151123/"
ptcov.dir <- my.dir %&% "genotypes/"
```

###pull height phenotype for GCTA
```{r}
norway <- read.csv(my.dir %&% "Norway_20150122/460platina110113 Helle.csv") %>% mutate(patno=as.character(Lopenr),C_HTCM=høyde2cm) %>% dplyr::select(patno,C_HTCM)
demo <- read.sas7bdat("/Volumes/dolan-lab/hwheeler/ThePlatinumStudy/GWAS/Datasets_from_Ryan_Cook_20151106/demographics.sas7bdat") %>% mutate(patno=as.character(patno)) %>% dplyr::select(patno,C_HTCM) #more ht phenotypes in this dataset
hist(demo$C_HTCM)
fam <- read.table(my.dir %&% "genotypes/gctaQC/N88_Recluster_TOP_20150911_FinalReport.postQC.fam")
colnames(fam) <- c("FID","IID","father","mother","sex","pheno")
patnoinfo <- read.table(my.dir %&% "genotypes/FID_IID_patno.txt",header=TRUE)
genosamples <- inner_join(fam,patnoinfo,by="FID")
demoComb <- rbind(demo,norway)
demoComb <- demoComb[!duplicated(demoComb$patno), ]
height <- left_join(genosamples,demoComb,by="patno")
heightpheno <- dplyr::select(height,FID,IID.x,C_HTCM)
dim(heightpheno)
hist(heightpheno$C_HTCM)
write.table(heightpheno,my.dir %&% "GCTA/height_cm.phen",quote=FALSE,row.names = FALSE, col.names = FALSE)
##make shuffled phenofile
set.seed(101)
shuffleht <- base::sample(heightpheno$C_HTCM)
shufpheno <- dplyr::select(heightpheno,FID,IID.x) %>% mutate(shuffleht=shuffleht)
write.table(shufpheno,my.dir %&% "GCTA/shuffled_height_cm.phen",quote=FALSE,row.names = FALSE, col.names = FALSE)
```
###GCTA results
- See 07_gcta.sh for code

#### Height phenotype

```
Source	Variance	SE
V(G)	33.573465	13.114136
V(e)	19.141336	12.920689
Vp	52.714800	2.476070
V(G)/Vp	0.636889	0.245407
logL	-2258.906
logL0	-2270.947
LRT	24.081
df	1
Pval	4.618e-07
n	910
```
####Shuffled height phenotype

```
Source	Variance	SE
V(G)	14.091383	16.237441
V(e)	39.849214	16.287943
Vp	53.940596	2.530541
V(G)/Vp	0.261239	0.300564
logL	-2270.443
logL0	-2270.947
LRT	1.008
df	1
Pval	0.1576
n	910
```

### Samples in QC'd genotypes that have exome data
- N88_Recluster_TOP_20150911_FinalReport.postPCA.euro.fam
```{r}
exomelist <- read.table("../RIKEN_genotyping/DSR_exomes_PATNO_kitNumber.txt",header=TRUE) 
colnames(exomelist)<-c("IID","kitNumber") #submitted exome list
completedexomes <- read.table("../RIKEN_genotyping/DSR_completed_dolan_pgrn_co_1_lookup_table.txt",header=TRUE,sep="\t") %>% mutate(kitNumber=UDF.Investigator.Sample.Name) %>% dplyr::select(Sample.Name,kitNumber)
exomes <- inner_join(exomelist,completedexomes,by="kitNumber") %>% mutate(exome="Yes",IID=as.character(IID))
dim(exomes)
fam <- read.table(my.dir %&% "genotypes/N88_Recluster_TOP_20150911_FinalReport.postPCA.euro.fam")
colnames(fam) <- c("FID","IID","father","mother","sex","pheno")
gtinfo <- left_join(fam,exomes,by="IID") %>% mutate(exome=ifelse(is.na(exome),"No","Yes")) %>% dplyr::select(FID,IID,kitNumber,Sample.Name,exome)
write.table(gtinfo,my.dir %&% "PtStudy_passed_genoQC_exome_info.txt",quote=F,row.names = F)
##how many have exomes?
table(gtinfo$exome)
```

### How many Norway genotyped samples have SCIN?
```{r}
demo <- read.sas7bdat("/Volumes/dolan-lab/hwheeler/ThePlatinumStudy/GWAS/Datasets_from_Ryan_Cook_20151106/demographics.sas7bdat") %>% mutate(patno=as.character(patno))
fam <- read.table(my.dir %&% "genotypes/N88_Recluster_TOP_20150911_FinalReport.fam")
colnames(fam) <- c("FID","IID","father","mother","sex","pheno")
patnoinfo <- read.table(my.dir %&% "genotypes/FID_IID_patno.txt",header=TRUE) %>% mutate(patno=as.character(patno))
genosamples <- inner_join(fam,patnoinfo,by="FID")
norgenosamples <- dplyr::filter(genosamples,grepl('T',IID.x))
##number Norway samples genotyped
dim(norgenosamples)[1]
norgenoscin <- inner_join(norgenosamples,demo,by="patno")
##number Norway samples genotyped with SCIN
dim(norgenoscin)[1]
##number Norway samples need SCIN for 
dim(norgenosamples)[1]-dim(norgenoscin)[1]

morescin <-read.spss('/Users/heather/GitHub/ThePlatinumStudy/Ototoxicity_modeling/Norway_audiometry_20150221//Travis_DNA_audiogram CHEM_195 21.02.15.sav',to.data.frame=T) %>% mutate(patno=as.character(lopenr))
norgenoscin2 <- inner_join(norgenosamples,morescin,by="patno")
dim(norgenoscin2)

##make list for Eileen J. of what we have and need
norgenoinfo <- left_join(norgenosamples,demo,by="patno") %>% mutate(lopenr=patno,IID=IID.x) %>% dplyr::select(lopenr,FID,IID,C_Q27ARSP)
aud <- read.sas7bdat("/Volumes/dolan-lab/hwheeler/ThePlatinumStudy/GWAS/Datasets_from_Ryan_Cook_20151106/audiometry.sas7bdat") %>% mutate(lopenr=as.character(patno)) %>% filter(patno<30000,C_Audtest==1,C_EARLR==1)
norgenoinfo <- left_join(norgenoinfo,aud,by="lopenr") 
a<- norgenoinfo %>% mutate(SCIN=ifelse(is.na(C_Q27ARSP),"No","Yes"), Audiometry=ifelse(is.na(C_EARLR),"No","Yes")) %>% dplyr::select(lopenr,FID,IID,SCIN,Audiometry)
write.table(a,file=my.dir %&% "Genotyped_Norway_samples_phenotype_info_2015-11-16.csv",sep=",",quote=F,row.names=F)
```
###Pull 4-8kHz audiometry geom mean phenotype for Norway samples
```{r}
aud <- read.sas7bdat("/Volumes/dolan-lab/hwheeler/ThePlatinumStudy/GWAS/Datasets_from_Ryan_Cook_20151106/audiometry.sas7bdat") %>% mutate(lopenr=as.character(patno)) %>% filter(patno<30000,C_Audtest==1)
##calc GM 4-8 kHz
airl<-filter(aud,C_EARLR==1)
airr<-filter(aud,C_EARLR==2)
###Pull db levels for 4-8 kHz
left <- airl %>% dplyr::select(C_CND4000, C_CND6000, C_CND8000)
right <- airr %>% dplyr::select(C_CND4000, C_CND6000, C_CND8000)
mean48 <- elementMean(list(left,right))
mean48<-replace(mean48,mean48<=1,1) ###set any db values <1 equal to 1, so can take the log
GM48 <- exp(rowMeans(log(mean48),na.rm=TRUE)) ##calc even if missing values
norGM48<- airl %>% dplyr::select(lopenr) %>% mutate(GM48nor=GM48)
norGM48phen <- inner_join(norGM48,norgenoinfo,by='lopenr') %>% mutate(ageaudio=Age_Audiogram)
rnGM48nor <- rntransform(norGM48phen$GM48nor)
norGM48phenfile <- norGM48phen %>% dplyr::select(FID,IID,ageaudio,GM48nor) %>% mutate(rnGM48nor=rnGM48nor)
write.table(norGM48phenfile, ptcov.dir %&% "Norway.phenofile",quote=F,row.names=F)
```

###Pull 4-12kHz audiometry geom mean phenotype
```{r}
fam <- read.table(my.dir %&% "genotypes/N88.forImputation.fam")
colnames(fam) <- c("FID","patno","father","mother","sex","pheno") 
fam <- fam %>% mutate(IID=as.character(patno))
aud <- read.sas7bdat(pheno.dir %&% "audiometry.sas7bdat")
ageataudio <- read.sas7bdat(pheno.dir %&% "demographics.sas7bdat") %>% mutate(IID=as.character(patno))
finaldosing <- read.sas7bdat(pheno.dir %&% "finaldosing.sas7bdat") %>% mutate(IID=as.character(patno)) %>% dplyr::select(IID,CumlCispdose,CumlBLEOdose)
##divide by ear
airl<-filter(aud,C_AUDTEST==1,C_EARLR==1)
airr<-filter(aud,C_AUDTEST==1,C_EARLR==2)
###Pull db levels for 4-12 kHz
left <- airl %>% dplyr::select(C_CND4000, C_CND6000, C_CND8000, C_CND10000, C_CND12000)
right <- airr %>% dplyr::select(C_CND4000, C_CND6000, C_CND8000, C_CND10000, C_CND12000)
mean412 <- elementMean(list(left,right))
mean412<-replace(mean412,mean412<=1,1) ###set any db values <1 equal to 1, so can take the log
GM412 <- exp(rowMeans(log(mean412),na.rm=TRUE)) ##calc even if missing values
patGM412<- airl %>% dplyr::select(patno) %>% mutate(GM412=GM412,IID=as.character(patno))
patGM412 <- left_join(fam,patGM412,by="IID")
patageaudio <- left_join(fam,ageataudio,by="IID") %>% dplyr::select(FID,IID,ageaudio)
patcov <- left_join(patageaudio,finaldosing,by="IID")
hist(patGM412$GM412)
hist(log(patGM412$GM412))
shapiro.test(log(patGM412$GM412)) #rank-normalize?

rnGM412 <- rntransform(patGM412$GM412)
hist(rnGM412)
patGM412 <- patGM412 %>% dplyr::select(FID,IID,GM412) %>% mutate(logGM412=log(GM412),rnGM412=rnGM412)

europcs <- read.table(my.dir %&% "genotypes/N88_euro_LDpruned_2015-11-18.evec",skip=1)
eupcdf <- europcs %>% rename(PC1=V2,PC2=V3,PC3=V4,PC4=V5,PC5=V6,PC6=V7,PC7=V8,PC8=V9,PC9=V10,PC10=V11)
patcov <- cbind(patcov,eupcdf[,2:11])

write.table(patGM412,file=my.dir %&% "genotypes/PtStudy.GM412.phenofile",quote=F,row.names = F,col.names = T)
write.table(patcov,file=my.dir %&% "genotypes/PtStudy.covfile",quote=F,row.names = F,col.names = T)

summary(lm(patGM412$rnGM412~patageaudio$ageaudio))$coef
summary(lm(patGM412$rnGM412~patageaudio$ageaudio))$r.squared

#number nonmissing rnGM412 phenotypes (TRUE):
table(is.na(patGM412$rnGM412)==FALSE)
#number nonmissing ageaudio (TRUE):
table(is.na(patcov$ageaudio)==FALSE)
#number nonmissing CumlCispdose (TRUE):
table(is.na(patcov$CumlCispdose)==FALSE)

#pull Pt Study only
poppatGM412 <- patGM412 %>% mutate(pop=ifelse(grepl("TC",IID),"Norway","PtStudy"))
table(poppatGM412$pop)
PtpatGM412 <- poppatGM412 %>% dplyr::filter(pop=="PtStudy")
#number nonmissing rnGM412 phenotypes (TRUE):
table(is.na(PtpatGM412$rnGM412)==FALSE)
table(is.na(PtpatGM412$rnGM412)==FALSE)/sum(table(is.na(PtpatGM412$rnGM412)==FALSE))
```

###Divide sample into quartiles based on Engdahl
```{r, quartiles, echo=TRUE, warning=FALSE,message=FALSE,fig.height=4, fig.width=5}
###Pull db levels for 4-8 kHz
left <- airl %>% dplyr::select(C_CND4000, C_CND6000, C_CND8000)
right <- airr %>% dplyr::select(C_CND4000, C_CND6000, C_CND8000)
mean468 <- elementMean(list(left,right))
mean468<-replace(mean468,mean468<=1,1) ###set any db values <1 equal to 1, so can take the log
AUC468 <- exp(rowMeans(log(mean468)))
patData <- airl %>% dplyr::select(patno) %>% mutate(AUC_4_8kHz=AUC468)
a <- left_join(patData,ageataudio,by="patno") 
patData <- a %>% dplyr::select(patno, ageaudio, AUC_4_8kHz)
###Define age groups per sample
group <- floor(patData$ageaudio) ##what I did previously, more pleasant graphic
for(i in 1:length(group)){
  if(is.na(group[i])){
    group[i] <- NA
  }
  else if(group[i] <= 29){
    group[i] <- 29
  }else if(group[i] > 29 && group[i] <= 39){
    group[i] <- 39
  }else if(group[i] > 39 && group[i] <= 49){
    group[i] <- 49
  }else if(group[i] > 49 && group[i] <= 59){
    group[i] <- 59
  }else if (group[i] > 59 && group[i] <= 69){
    group[i] <- 69
  }
}
patData <- patData %>% mutate(GROUP=group)
##4-8kHz normal AUCs calculated in Engdahl_def_quartiles.xlsx
engData <- read.table('~/GitHub/ThePlatinumStudy/Ototoxicity_modeling/Engdahl_quartile_AUCs_4-8kHz.txt',header=TRUE)
aucData <- left_join(patData,engData,by="GROUP")
aucData <- aucData[complete.cases(aucData),]

###calculate which quartile
quartile=aucData$AUC_4_8kHz
for(i in 1:dim(aucData)[1]){
  if(aucData$AUC_4_8kHz[i] < aucData$Q1[i]){
    quartile[i] <- 1
  }
  else if(aucData$AUC_4_8kHz[i] < aucData$Q2[i]){
    quartile[i] <- 2
  }
  else if(aucData$AUC_4_8kHz[i] < aucData$Q3[i]){
    quartile[i] <- 3
  }else{
    quartile[i] <- 4
  }
}

aucData <- mutate(aucData,QUARTILE=quartile,IID=as.character(patno))
phenCombo <- left_join(patGM412,aucData,by="IID")

table(aucData$QUARTILE)
sum(table(aucData$QUARTILE))
table(aucData$QUARTILE)/sum(table(aucData$QUARTILE))
```

###Pull EORTC CIPN20: 8 sensory items (no hearing item)
```{r}
fam <- read.table(my.dir %&% "genotypes/N88.forImputation.fam")
colnames(fam) <- c("FID","patno","father","mother","sex","pheno") 
fam <- fam %>% mutate(IID=as.character(patno))
demog <- read.sas7bdat(pheno.dir %&% "demographics.sas7bdat")
cipn8sens <- demog %>% dplyr::select(patno, agediagnosis, C_Q15ARSP, C_Q15BRSP, C_Q15CRSP, C_Q15DRSP, C_Q15ERSP, C_Q15FRSP, C_Q15IRSP, C_Q15JRSP) %>% rename(Tingling_hands=C_Q15ARSP, Tingling_feet=C_Q15BRSP, Numbness_hands=C_Q15CRSP, Numbness_feet=C_Q15DRSP, Shooting_pain_hands=C_Q15ERSP, Shooting_pain_feet=C_Q15FRSP, Problems_walking_feeling_ground=C_Q15IRSP, Difficulty_hot_cold_water=C_Q15JRSP)

##Plot responses
a<-cipn8sens
ga <- gather(a,question,response,-patno, -agediagnosis) 
ggplot(ga,aes(x=factor(response,labels=c("not\nat all","a little","quite\na bit","very\nmuch","NA")), fill=factor(response)))+geom_bar()+facet_wrap(~question)+xlab("Response")+geom_text(aes(y = (..count..),label = paste(round(ifelse((..count..)==0,"",8*((..count..)/sum(..count..)))*100,1),"%",sep="")), stat="count",vjust=-0.25,size=4)+guides(fill=FALSE)+theme_bw()+ylab("Number of patients")+coord_cartesian(ylim=c(-20,900))

##Calculate Cronbach's alpha
CIPN20_8sens <- cipn8sens[,3:10]
psych::alpha(CIPN20_8sens)$total

cipn8m <- dplyr::select(cipn8sens,patno,agediagnosis) %>% mutate(IID=as.character(patno),meanCIPN8=rowMeans(CIPN20_8sens,na.rm=TRUE))
hist(cipn8m$meanCIPN8)
table(cipn8m$meanCIPN8==0)
table(ceiling(cipn8m$meanCIPN8))
table(ceiling(cipn8m$meanCIPN8))/sum(table(ceiling(cipn8m$meanCIPN8)))
hist(ceiling(cipn8m$meanCIPN8))
##ordCIPN8 groups: meanCIPN8: 0, 0<x<=1, 1<x<=2, 2<x<=3
##biCIPN8: meanCIPN8=0, >0
##exCIPN8: meanCIPN8=0 vs 2or3
##ord3CIPN8 groups: meanCIPN8: 0, 0<x<=1, 1<x<=3
cipn8m <- dplyr::select(cipn8m,IID,agediagnosis,meanCIPN8) %>% mutate(ordCIPN8=ceiling(meanCIPN8),biCIPN8=ifelse(meanCIPN8==0,0,1)) %>% mutate(ord3CIPN8=ifelse(ordCIPN8==3,2,ordCIPN8))
exCIPN8 <- ifelse(cipn8m$ordCIPN8==1,NA,cipn8m$ordCIPN8)
exCIPN8 <- ifelse(exCIPN8>=2,1,exCIPN8)
cipn8m <- cipn8m %>% mutate(exCIPN8=exCIPN8)
table(cipn8m$exCIPN8)
table(cipn8m$ord3CIPN8)
hist(cipn8m$ord3CIPN8)
head(cipn8m)
dim(cipn8m)

patcipn <- left_join(fam,cipn8m,by="IID") %>% dplyr::select(FID,IID,meanCIPN8,ordCIPN8,biCIPN8,exCIPN8,ord3CIPN8,agediagnosis)

write.table(patcipn,file=my.dir %&% "genotypes/PtStudy.CIPN8.phenofile",quote=F,row.names = F,col.names = T)

nona.patcipn <- patcipn[is.na(patcipn$ordCIPN8)==FALSE,]
ggplot(nona.patcipn,aes(x=factor(ordCIPN8,labels=c("not\nat all","a little","quite\na bit","very\nmuch")), fill=factor(ordCIPN8)))+geom_bar()+xlab("Response")+geom_text(aes(y = (..count..),label = paste(round(ifelse((..count..)==0,"",((..count..)/sum(..count..)))*100,1),"%",sep="")), stat="count",vjust=-0.25,size=4)+guides(fill=FALSE)+theme_bw()+ylab("Number of patients")+ggtitle('ordCIPN8 (ceiling of mean CIPN8)')

ggplot(nona.patcipn,aes(x=factor(ord3CIPN8,labels=c("not\nat all","a little","quite a bit or\nvery much")), fill=factor(ord3CIPN8)))+geom_bar()+xlab("Response")+geom_text(aes(y = (..count..),label = paste(round(ifelse((..count..)==0,"",((..count..)/sum(..count..)))*100,1),"%",sep="")), stat="count",vjust=-0.25,size=4)+guides(fill=FALSE)+theme_bw()+ylab("Number of patients")+ggtitle('ord3CIPN8 (ceiling of mean CIPN8, combine 2+3)')

summary(lm(patcipn$agediagnosis~patcipn$ordCIPN8))$coef
summary(lm(patcipn$agediagnosis~patcipn$ordCIPN8))$r.squared

#number nonmissing ordCIPN8 phenotypes (TRUE):
table(is.na(patcipn$ordCIPN8)==FALSE)
#number nonmissing exCIPN8 phenotypes (TRUE):
table(is.na(patcipn$exCIPN8)==FALSE)

phenCombo <- left_join(phenCombo,patcipn,by="IID")
phenCombo <- left_join(phenCombo,patcov,by="IID") 
phenEric <- phenCombo %>% dplyr::select(FID.x, IID, GM412, rnGM412, logGM412, QUARTILE ,meanCIPN8,ordCIPN8,biCIPN8,ord3CIPN8,exCIPN8,ageaudio.x,agediagnosis,CumlCispdose,CumlBLEOdose) %>% dplyr::rename(FID=FID.x,quartile=QUARTILE,ageaudio=ageaudio.x)
phentable <- phenEric %>% mutate(CumlBLEOdose=ifelse(is.nan(CumlBLEOdose),NA,CumlBLEOdose)) #should any CumlBLEOdose == 0? Check before Raynaud's GWAS
```

test run:  
`plink --bfile N88.forImputation --1 --pheno PtStudy.CIPN8.phenofile --pheno-name biCIPN8 --logistic --adjust --covar PtStudy.covfile --out PtStudy.preImpute.2015-11-24.biCIPN8ageaudio.Cispdose`

`plink --bfile N88.forImputation --1 --pheno PtStudy.CIPN8.phenofile --pheno-name exCIPN8 --logistic --adjust --covar PtStudy.covfile --out PtStudy.preImpute.2015-11-24.exCIPN8ageaudio.Cispdose`

`plink --bfile N88.forImputation --1 --pheno PtStudy.CIPN8.phenofile --pheno-name biCIPN8 --logistic --adjust --covar PtStudy.covfile --covar-name ageaudio --out PtStudy.preImpute.2015-11-24.biCIPN8ageaudio`

`plink --bfile N88.forImputation --1 --pheno PtStudy.CIPN8.phenofile --pheno-name exCIPN8 --logistic --adjust --covar PtStudy.covfile --covar-name ageaudio --out PtStudy.preImpute.2015-11-24.exCIPN8ageaudio`

####TO DO: code for ordinal regression phenotypes-->done, see 14_ordreg_GWAS.R

###Does ordCIPN8 associate with agediagnosis? CumlCispdose?
- only agediagnosis is signficant:
```{r cov}
phenofile <- ptcov.dir %&% "PtStudy.phenofile"
covfile <- ptcov.dir %&% "PtStudy.covfile"
pheno <- read.table(phenofile,header=T)
cov <- read.table(covfile,header=T) %>% dplyr::select(FID,starts_with("PC"))
pheno <- left_join(pheno,cov,by="FID")

#ordCIPN8 ~ agediagnosis
m<-polr(factor(pheno$ordCIPN8) ~ pheno$agediagnosis, Hess=TRUE)
p <- pnorm(abs(summary(m)$coefficients[, "t value"]), lower.tail = FALSE) * 2
or <- exp(summary(m)$coefficients[,"Value"])
ci <- exp(confint.default(m))
colnames(ci) <- c("lCI","uCI")
res <- cbind(summary(m)$coefficients[,],"OR"=or,"pval"=p)[1,]
res <- c(res, ci[1,])
res

#ordCIPN8 ~ CumlCispdose
m<-polr(factor(pheno$ordCIPN8) ~ pheno$CumlCispdose, Hess=TRUE)
p <- pnorm(abs(summary(m)$coefficients[, "t value"]), lower.tail = FALSE) * 2
or <- exp(summary(m)$coefficients[,"Value"])
ci <- exp(confint.default(m))
colnames(ci) <- c("lCI","uCI")
res <- cbind(summary(m)$coefficients[,],"OR"=or,"pval"=p)[1,]
res <- c(res, ci[1,])
res

#ordCIPN8 ~ agediagnosis + CumlCispdose
m<-polr(factor(pheno$ordCIPN8) ~ pheno$agediagnosis + pheno$CumlCispdose, Hess=TRUE)
p <- pnorm(abs(summary(m)$coefficients[, "t value"]), lower.tail = FALSE) * 2
or <- exp(summary(m)$coefficients[,"Value"])
ci <- exp(confint.default(m))
colnames(ci) <- c("lCI","uCI")
res <- cbind(summary(m)$coefficients[,],"OR"=or,"pval"=p)
res <- cbind(res[1:2,],ci)
res

#ordCIPN8 ~ agediagnosis + 10 PCs
covmat <- dplyr::select(pheno,agediagnosis,starts_with("PC"))
covmat <- as.matrix(covmat)
m<-polr(factor(pheno$ordCIPN8) ~ covmat, Hess=TRUE)
p <- pnorm(abs(summary(m)$coefficients[, "t value"]), lower.tail = FALSE) * 2
or <- exp(summary(m)$coefficients[,"Value"])
ci <- exp(confint.default(m))
colnames(ci) <- c("lCI","uCI")
res <- cbind(summary(m)$coefficients[,],"OR"=or,"pval"=p)
res <- cbind(res[1:dim(covmat)[2],], ci)
res

#ordCIPN8 ~ agediagnosis + CumlCispdose + 10 PCs
covmat <- dplyr::select(pheno,agediagnosis,CumlCispdose,starts_with("PC"))
covmat <- as.matrix(covmat)
m<-polr(factor(pheno$ordCIPN8) ~ covmat, Hess=TRUE)
p <- pnorm(abs(summary(m)$coefficients[, "t value"]), lower.tail = FALSE) * 2
or <- exp(summary(m)$coefficients[,"Value"])
ci <- exp(confint.default(m))
colnames(ci) <- c("lCI","uCI")
res <- cbind(summary(m)$coefficients[,],"OR"=or,"pval"=p)
res <- cbind(res[1:dim(covmat)[2],], ci)
res
```

###Pull SCIN
- Tinnitus: ordTin (C_Q27ERSP)
- Reduced Hearing: ordHL (C_Q27FRSP)
- Raynaud's: ordRay (mean C_Q27CRSP & C_Q27DRSP, take ceiling)
- also pull CumlBLEOdose
- add to PtStudy.phenofile
```{r}
fam <- read.table(my.dir %&% "genotypes/N88.forImputation.fam")
colnames(fam) <- c("FID","patno","father","mother","sex","pheno") 
fam <- fam %>% mutate(IID=as.character(patno))
demog <- read.sas7bdat(pheno.dir %&% "demographics.sas7bdat")
scin <- demog %>% dplyr::select(patno,C_Q27ARSP,C_Q27BRSP,C_Q27CRSP,C_Q27DRSP,C_Q27ERSP,C_Q27FRSP) %>% rename(Pain_Tingling_Feet=C_Q27ARSP,Pain_Tingling_Hands=C_Q27BRSP,White_Cold_Hands=C_Q27CRSP,White_Cold_Feet=C_Q27DRSP,ordTin=C_Q27ERSP,ordHL=C_Q27FRSP)
a<-scin
dim(a)
##Plot responses
ga <- gather(a,question,response,-patno) 
ga <- ga %>% mutate(type=ifelse(question=="Pain_Tingling_Feet" | question=="Pain_Tingling_Hands","parathesias",question)) %>% mutate(type=ifelse(type<=4,"Raynaud",type)) %>% mutate(type=ifelse(type<=6,"ototoxicity",type))
ggplot(ga,aes(x=factor(response,labels=c("not\nat all","a little","quite\na bit","very\nmuch","NA")),fill=type))+geom_bar()+facet_wrap(~question)+xlab("Response")+geom_text(aes(y = (..count..),label = paste(round(ifelse((..count..)==0,"",6*((..count..)/sum(..count..)))*100,1),"%",sep="")), stat="count",vjust=-0.25,size=4)+guides(fill=FALSE)+theme_bw()+ylab("Number of patients")+coord_cartesian(ylim=c(-20,750))

ray <- dplyr::select(scin,White_Cold_Hands,White_Cold_Feet)
ordRay <- ceiling(rowMeans(ray,na.rm=TRUE))
hist(ordRay)
table(ordRay)/sum(table(ordRay))
morepheno <- dplyr::select(scin,patno,ordTin,ordHL) %>% mutate(IID=as.character(patno),ordRay=ordRay)
hist(morepheno$ordTin)
table(morepheno$ordTin)/sum(table(morepheno$ordTin))
hist(morepheno$ordHL)
table(morepheno$ordHL)/sum(table(morepheno$ordHL))

phentable2<-left_join(phentable,morepheno,by="IID")
#write.table(phentable2,file=my.dir %&% "genotypes/PtStudy.phenofile",quote=F,row.names = F,col.names = T) --> this file is now PtStudy.phenofile.20160119
```

###Do ordTin and ordHL associate with agediagnosis? CumlCispdose?
- only agediagnosis is signficant:
```{r}
phenofile <- ptcov.dir %&% "PtStudy.phenofile"
covfile <- ptcov.dir %&% "PtStudy.covfile"
pheno <- read.table(phenofile,header=T)
cov <- read.table(covfile,header=T) %>% dplyr::select(FID,starts_with("PC"))
pheno <- left_join(pheno,cov,by="FID")

#ordTin ~ agediagnosis
m<-polr(factor(pheno$ordTin) ~ pheno$agediagnosis, Hess=TRUE)
p <- pnorm(abs(summary(m)$coefficients[, "t value"]), lower.tail = FALSE) * 2
or <- exp(summary(m)$coefficients[,"Value"])
ci <- exp(confint.default(m))
colnames(ci) <- c("lCI","uCI")
res <- cbind(summary(m)$coefficients[,],"OR"=or,"pval"=p)[1,]
res <- c(res, ci[1,])
res

#ordTin ~ CumlCispdose
m<-polr(factor(pheno$ordTin) ~ pheno$CumlCispdose, Hess=TRUE)
p <- pnorm(abs(summary(m)$coefficients[, "t value"]), lower.tail = FALSE) * 2
or <- exp(summary(m)$coefficients[,"Value"])
ci <- exp(confint.default(m))
colnames(ci) <- c("lCI","uCI")
res <- cbind(summary(m)$coefficients[,],"OR"=or,"pval"=p)[1,]
res <- c(res, ci[1,])
res

#ordTin ~ agediagnosis + CumlCispdose
m<-polr(factor(pheno$ordTin) ~ pheno$agediagnosis + pheno$CumlCispdose, Hess=TRUE)
p <- pnorm(abs(summary(m)$coefficients[, "t value"]), lower.tail = FALSE) * 2
or <- exp(summary(m)$coefficients[,"Value"])
ci <- exp(confint.default(m))
colnames(ci) <- c("lCI","uCI")
res <- cbind(summary(m)$coefficients[,],"OR"=or,"pval"=p)
res <- cbind(res[1:2,],ci)
res

#ordTin ~ agediagnosis + 10 PCs
covmat <- dplyr::select(pheno,agediagnosis,starts_with("PC"))
covmat <- as.matrix(covmat)
m<-polr(factor(pheno$ordTin) ~ covmat, Hess=TRUE)
p <- pnorm(abs(summary(m)$coefficients[, "t value"]), lower.tail = FALSE) * 2
or <- exp(summary(m)$coefficients[,"Value"])
ci <- exp(confint.default(m))
colnames(ci) <- c("lCI","uCI")
res <- cbind(summary(m)$coefficients[,],"OR"=or,"pval"=p)
res <- cbind(res[1:dim(covmat)[2],], ci)
res

#ordTin ~ agediagnosis + CumlCispdose + 10 PCs
covmat <- dplyr::select(pheno,agediagnosis,CumlCispdose,starts_with("PC"))
covmat <- as.matrix(covmat)
m<-polr(factor(pheno$ordTin) ~ covmat, Hess=TRUE)
p <- pnorm(abs(summary(m)$coefficients[, "t value"]), lower.tail = FALSE) * 2
or <- exp(summary(m)$coefficients[,"Value"])
ci <- exp(confint.default(m))
colnames(ci) <- c("lCI","uCI")
res <- cbind(summary(m)$coefficients[,],"OR"=or,"pval"=p)
res <- cbind(res[1:dim(covmat)[2],], ci)
res

#ordHL ~ agediagnosis
m<-polr(factor(pheno$ordHL) ~ pheno$agediagnosis, Hess=TRUE)
p <- pnorm(abs(summary(m)$coefficients[, "t value"]), lower.tail = FALSE) * 2
or <- exp(summary(m)$coefficients[,"Value"])
ci <- exp(confint.default(m))
colnames(ci) <- c("lCI","uCI")
res <- cbind(summary(m)$coefficients[,],"OR"=or,"pval"=p)[1,]
res <- c(res, ci[1,])
res

#ordHL ~ CumlCispdose
m<-polr(factor(pheno$ordHL) ~ pheno$CumlCispdose, Hess=TRUE)
p <- pnorm(abs(summary(m)$coefficients[, "t value"]), lower.tail = FALSE) * 2
or <- exp(summary(m)$coefficients[,"Value"])
ci <- exp(confint.default(m))
colnames(ci) <- c("lCI","uCI")
res <- cbind(summary(m)$coefficients[,],"OR"=or,"pval"=p)[1,]
res <- c(res, ci[1,])
res

#ordHL ~ agediagnosis + CumlCispdose
m<-polr(factor(pheno$ordHL) ~ pheno$agediagnosis + pheno$CumlCispdose, Hess=TRUE)
p <- pnorm(abs(summary(m)$coefficients[, "t value"]), lower.tail = FALSE) * 2
or <- exp(summary(m)$coefficients[,"Value"])
ci <- exp(confint.default(m))
colnames(ci) <- c("lCI","uCI")
res <- cbind(summary(m)$coefficients[,],"OR"=or,"pval"=p)
res <- cbind(res[1:2,],ci)
res

#ordHL ~ agediagnosis + 10 PCs
covmat <- dplyr::select(pheno,agediagnosis,starts_with("PC"))
covmat <- as.matrix(covmat)
m<-polr(factor(pheno$ordHL) ~ covmat, Hess=TRUE)
p <- pnorm(abs(summary(m)$coefficients[, "t value"]), lower.tail = FALSE) * 2
or <- exp(summary(m)$coefficients[,"Value"])
ci <- exp(confint.default(m))
colnames(ci) <- c("lCI","uCI")
res <- cbind(summary(m)$coefficients[,],"OR"=or,"pval"=p)
res <- cbind(res[1:dim(covmat)[2],], ci)
res

#ordHL ~ agediagnosis + CumlCispdose + 10 PCs
covmat <- dplyr::select(pheno,agediagnosis,CumlCispdose,starts_with("PC"))
covmat <- as.matrix(covmat)
m<-polr(factor(pheno$ordHL) ~ covmat, Hess=TRUE)
p <- pnorm(abs(summary(m)$coefficients[, "t value"]), lower.tail = FALSE) * 2
or <- exp(summary(m)$coefficients[,"Value"])
ci <- exp(confint.default(m))
colnames(ci) <- c("lCI","uCI")
res <- cbind(summary(m)$coefficients[,],"OR"=or,"pval"=p)
res <- cbind(res[1:dim(covmat)[2],], ci)
res
```

###Does ordRay associate with agediagnosis? CumlBLEOdose?
- only agediagnosis is signficant:
```{r}
phenofile <- ptcov.dir %&% "PtStudy.phenofile"
covfile <- ptcov.dir %&% "PtStudy.covfile"
pheno <- read.table(phenofile,header=T)
cov <- read.table(covfile,header=T) %>% dplyr::select(FID,starts_with("PC"))
pheno <- left_join(pheno,cov,by="FID")

#ordRay ~ agediagnosis
m<-polr(factor(pheno$ordRay) ~ pheno$agediagnosis, Hess=TRUE)
p <- pnorm(abs(summary(m)$coefficients[, "t value"]), lower.tail = FALSE) * 2
or <- exp(summary(m)$coefficients[,"Value"])
ci <- exp(confint.default(m))
colnames(ci) <- c("lCI","uCI")
res <- cbind(summary(m)$coefficients[,],"OR"=or,"pval"=p)[1,]
res <- c(res, ci[1,])
res

#ordRay ~ CumlBLEOdose
m<-polr(factor(pheno$ordRay) ~ pheno$CumlBLEOdose, Hess=TRUE)
p <- pnorm(abs(summary(m)$coefficients[, "t value"]), lower.tail = FALSE) * 2
or <- exp(summary(m)$coefficients[,"Value"])
ci <- exp(confint.default(m))
colnames(ci) <- c("lCI","uCI")
res <- cbind(summary(m)$coefficients[,],"OR"=or,"pval"=p)[1,]
res <- c(res, ci[1,])
res

#ordRay ~ agediagnosis + CumlBLEOdose
m<-polr(factor(pheno$ordRay) ~ pheno$agediagnosis + pheno$CumlBLEOdose, Hess=TRUE)
p <- pnorm(abs(summary(m)$coefficients[, "t value"]), lower.tail = FALSE) * 2
or <- exp(summary(m)$coefficients[,"Value"])
ci <- exp(confint.default(m))
colnames(ci) <- c("lCI","uCI")
res <- cbind(summary(m)$coefficients[,],"OR"=or,"pval"=p)
res <- cbind(res[1:2,],ci)
res

#ordRay ~ agediagnosis + 10 PCs
covmat <- dplyr::select(pheno,agediagnosis,starts_with("PC"))
covmat <- as.matrix(covmat)
m<-polr(factor(pheno$ordRay) ~ covmat, Hess=TRUE)
p <- pnorm(abs(summary(m)$coefficients[, "t value"]), lower.tail = FALSE) * 2
or <- exp(summary(m)$coefficients[,"Value"])
ci <- exp(confint.default(m))
colnames(ci) <- c("lCI","uCI")
res <- cbind(summary(m)$coefficients[,],"OR"=or,"pval"=p)
res <- cbind(res[1:dim(covmat)[2],], ci)
res

#ordRay ~ agediagnosis + CumlBLEOdose + 10 PCs
covmat <- dplyr::select(pheno,agediagnosis,CumlBLEOdose,starts_with("PC"))
covmat <- as.matrix(covmat)
m<-polr(factor(pheno$ordRay) ~ covmat, Hess=TRUE)
p <- pnorm(abs(summary(m)$coefficients[, "t value"]), lower.tail = FALSE) * 2
or <- exp(summary(m)$coefficients[,"Value"])
ci <- exp(confint.default(m))
colnames(ci) <- c("lCI","uCI")
res <- cbind(summary(m)$coefficients[,],"OR"=or,"pval"=p)
res <- cbind(res[1:dim(covmat)[2],], ci)
res
```