---
title: "06_make_phenofiles"
author: "Heather E. Wheeler"
date: "`r Sys.time()`"
output: html_document
---

###Load packages, functions, paths
```{r load,message=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(sas7bdat) ##read SAS files
library(foreign) ##read SPSS files, doesn't work with Fossa file, ask Ryan to convert to txt
library(psych)
library(MASS)
"%&%" = function(a,b) paste(a,b,sep="")
elementMean <- function(my.list) { 
  arr <- array(unlist(my.list),c(dim(my.list[[1]])[1],dim(my.list[[1]])[2],length(my.list)))
  rowMeans( arr , na.rm=TRUE, dims = 2 )
}
source("/Volumes/dolan-lab/hwheeler/ThePlatinumStudy/CIPN_modeling/ztransform.R")
source("/Volumes/dolan-lab/hwheeler/ThePlatinumStudy/CIPN_modeling/rntransform.R")
my.dir = "/Volumes/dolan-lab/hwheeler/ThePlatinumStudy/GWAS/"
pca.dir = "/Volumes/dolan-lab/hwheeler/ThePlatinumStudy/GWAS/PCA/"
pheno.dir = "/Volumes/dolan-lab/hwheeler/ThePlatinumStudy/GWAS/Datasets_from_Ryan_Cook_20160219/"
ptcov.dir <- my.dir %&% "genotypes/"
```


###Pull 4-12kHz audiometry geom mean phenotype
```{r}
fam <- read.table(my.dir %&% "genotypes/N88.forImputation.fam")
colnames(fam) <- c("FID","patno","father","mother","sex","pheno") 
fam <- fam %>% mutate(IID=as.character(patno))
aud <- read.sas7bdat(pheno.dir %&% "audiometry.sas7bdat")
ageataudio <- read.sas7bdat(pheno.dir %&% "demographics.sas7bdat") %>% mutate(IID=as.character(patno))
finaldosing <- read.sas7bdat(pheno.dir %&% "finaldosing.sas7bdat") %>% mutate(IID=as.character(PATNO)) %>% dplyr::select(IID,CumlCispdose,CumlBLEOdose)
##divide by ear
airl<-filter(aud,C_AUDTEST==1,C_EARLR==1)
airr<-filter(aud,C_AUDTEST==1,C_EARLR==2)
###Pull db levels for 4-12 kHz
left <- airl %>% dplyr::select(C_CND4000, C_CND6000, C_CND8000, C_CND10000, C_CND12000)
right <- airr %>% dplyr::select(C_CND4000, C_CND6000, C_CND8000, C_CND10000, C_CND12000)
mean412 <- elementMean(list(left,right))
mean412<-replace(mean412,mean412<=1,1) ###set any db values <1 equal to 1, so can take the log
GM412 <- exp(rowMeans(log(mean412),na.rm=TRUE)) ##calc even if missing values
patGM412<- airl %>% dplyr::select(patno) %>% mutate(GM412=GM412,IID=as.character(patno))
patGM412 <- left_join(fam,patGM412,by="IID")
patageaudio <- left_join(fam,ageataudio,by="IID") %>% dplyr::select(FID,IID,ageaudio)
patcov <- left_join(patageaudio,finaldosing,by="IID")
hist(patGM412$GM412)
hist(log(patGM412$GM412))
shapiro.test(log(patGM412$GM412)) #rank-normalize?

rnGM412 <- rntransform(patGM412$GM412)
hist(rnGM412)
patGM412 <- patGM412 %>% dplyr::select(FID,IID,GM412) %>% mutate(logGM412=log(GM412),rnGM412=rnGM412)

europcs <- read.table(my.dir %&% "genotypes/N88_euro_LDpruned_2015-11-18.evec",skip=1)
eupcdf <- europcs %>% rename(PC1=V2,PC2=V3,PC3=V4,PC4=V5,PC5=V6,PC6=V7,PC7=V8,PC8=V9,PC9=V10,PC10=V11)
patcov <- cbind(patcov,eupcdf[,2:11])

write.table(patGM412,file=my.dir %&% "genotypes/PtStudy.GM412.phenofile",quote=F,row.names = F,col.names = T)
write.table(patcov,file=my.dir %&% "genotypes/PtStudy.covfile",quote=F,row.names = F,col.names = T)

summary(lm(patGM412$rnGM412~patageaudio$ageaudio))$coef
summary(lm(patGM412$rnGM412~patageaudio$ageaudio))$r.squared

#number nonmissing rnGM412 phenotypes (TRUE):
table(is.na(patGM412$rnGM412)==FALSE)
#number nonmissing ageaudio (TRUE):
table(is.na(patcov$ageaudio)==FALSE)
#number nonmissing CumlCispdose (TRUE):
table(is.na(patcov$CumlCispdose)==FALSE)

#pull Pt Study only
poppatGM412 <- patGM412 %>% mutate(pop=ifelse(grepl("TC",IID),"Norway","PtStudy"))
table(poppatGM412$pop)
PtpatGM412 <- poppatGM412 %>% dplyr::filter(pop=="PtStudy")
#number nonmissing rnGM412 phenotypes (TRUE):
table(is.na(PtpatGM412$rnGM412)==FALSE)
table(is.na(PtpatGM412$rnGM412)==FALSE)/sum(table(is.na(PtpatGM412$rnGM412)==FALSE))
```

###Divide sample into quartiles based on Engdahl
```{r, quartiles, echo=TRUE, warning=FALSE,message=FALSE,fig.height=4, fig.width=5}
###Pull db levels for 4-8 kHz
left <- airl %>% dplyr::select(C_CND4000, C_CND6000, C_CND8000)
right <- airr %>% dplyr::select(C_CND4000, C_CND6000, C_CND8000)
mean468 <- elementMean(list(left,right))
mean468<-replace(mean468,mean468<=1,1) ###set any db values <1 equal to 1, so can take the log
AUC468 <- exp(rowMeans(log(mean468)))
patData <- airl %>% dplyr::select(patno) %>% mutate(AUC_4_8kHz=AUC468)
a <- left_join(patData,ageataudio,by="patno") 
patData <- a %>% dplyr::select(patno, ageaudio, AUC_4_8kHz)
###Define age groups per sample
group <- floor(patData$ageaudio) ##what I did previously, more pleasant graphic
for(i in 1:length(group)){
  if(is.na(group[i])){
    group[i] <- NA
  }
  else if(group[i] <= 29){
    group[i] <- 29
  }else if(group[i] > 29 && group[i] <= 39){
    group[i] <- 39
  }else if(group[i] > 39 && group[i] <= 49){
    group[i] <- 49
  }else if(group[i] > 49 && group[i] <= 59){
    group[i] <- 59
  }else if (group[i] > 59 && group[i] <= 69){
    group[i] <- 69
  }
}
patData <- patData %>% mutate(GROUP=group)
##4-8kHz normal AUCs calculated in Engdahl_def_quartiles.xlsx
engData <- read.table('~/GitHub/ThePlatinumStudy/Ototoxicity_modeling/Engdahl_quartile_AUCs_4-8kHz.txt',header=TRUE)
aucData <- left_join(patData,engData,by="GROUP")
aucData <- aucData[complete.cases(aucData),]

###calculate which quartile
quartile=aucData$AUC_4_8kHz
for(i in 1:dim(aucData)[1]){
  if(aucData$AUC_4_8kHz[i] < aucData$Q1[i]){
    quartile[i] <- 1
  }
  else if(aucData$AUC_4_8kHz[i] < aucData$Q2[i]){
    quartile[i] <- 2
  }
  else if(aucData$AUC_4_8kHz[i] < aucData$Q3[i]){
    quartile[i] <- 3
  }else{
    quartile[i] <- 4
  }
}

aucData <- mutate(aucData,QUARTILE=quartile,IID=as.character(patno))
phenCombo <- left_join(patGM412,aucData,by="IID")

table(aucData$QUARTILE)
sum(table(aucData$QUARTILE))
table(aucData$QUARTILE)/sum(table(aucData$QUARTILE))
```

###Pull EORTC CIPN20: 8 sensory items (no hearing item)
```{r}
fam <- read.table(my.dir %&% "genotypes/N88.forImputation.fam")
colnames(fam) <- c("FID","patno","father","mother","sex","pheno") 
fam <- fam %>% mutate(IID=as.character(patno))
demog <- read.sas7bdat(pheno.dir %&% "demographics.sas7bdat")
cipn8sens <- demog %>% dplyr::select(patno, agediagnosis, C_Q15ARSP, C_Q15BRSP, C_Q15CRSP, C_Q15DRSP, C_Q15ERSP, C_Q15FRSP, C_Q15IRSP, C_Q15JRSP) %>% rename(Tingling_hands=C_Q15ARSP, Tingling_feet=C_Q15BRSP, Numbness_hands=C_Q15CRSP, Numbness_feet=C_Q15DRSP, Shooting_pain_hands=C_Q15ERSP, Shooting_pain_feet=C_Q15FRSP, Problems_walking_feeling_ground=C_Q15IRSP, Difficulty_hot_cold_water=C_Q15JRSP)

##Calculate Cronbach's alpha
CIPN20_8sens <- cipn8sens[,3:10]
psych::alpha(CIPN20_8sens)$total

cipn8m <- dplyr::select(cipn8sens,patno,agediagnosis) %>% mutate(IID=as.character(patno),meanCIPN8=rowMeans(CIPN20_8sens,na.rm=TRUE))
hist(cipn8m$meanCIPN8)
table(cipn8m$meanCIPN8==0)
table(ceiling(cipn8m$meanCIPN8))
table(ceiling(cipn8m$meanCIPN8))/sum(table(ceiling(cipn8m$meanCIPN8)))
hist(ceiling(cipn8m$meanCIPN8))
##ordCIPN8 groups: meanCIPN8: 0, 0<x<=1, 1<x<=2, 2<x<=3
##biCIPN8: meanCIPN8=0, >0
##exCIPN8: meanCIPN8=0 vs 2or3
##ord3CIPN8 groups: meanCIPN8: 0, 0<x<=1, 1<x<=3
cipn8m <- dplyr::select(cipn8m,IID,agediagnosis,meanCIPN8) %>% mutate(ordCIPN8=ceiling(meanCIPN8),biCIPN8=ifelse(meanCIPN8==0,0,1)) %>% mutate(ord3CIPN8=ifelse(ordCIPN8==3,2,ordCIPN8))
exCIPN8 <- ifelse(cipn8m$ordCIPN8==1,NA,cipn8m$ordCIPN8)
exCIPN8 <- ifelse(exCIPN8>=2,1,exCIPN8)
cipn8m <- cipn8m %>% mutate(exCIPN8=exCIPN8)
table(cipn8m$exCIPN8)
table(cipn8m$ord3CIPN8)
hist(cipn8m$ord3CIPN8)
head(cipn8m)
dim(cipn8m)

patcipn <- left_join(fam,cipn8m,by="IID") %>% dplyr::select(FID,IID,meanCIPN8,ordCIPN8,biCIPN8,exCIPN8,ord3CIPN8,agediagnosis)

write.table(patcipn,file=my.dir %&% "genotypes/PtStudy.CIPN8.phenofile",quote=F,row.names = F,col.names = T)


summary(lm(patcipn$agediagnosis~patcipn$ordCIPN8))$coef
summary(lm(patcipn$agediagnosis~patcipn$ordCIPN8))$r.squared

#number nonmissing ordCIPN8 phenotypes (TRUE):
table(is.na(patcipn$ordCIPN8)==FALSE)
#number nonmissing exCIPN8 phenotypes (TRUE):
table(is.na(patcipn$exCIPN8)==FALSE)

phenCombo <- left_join(phenCombo,patcipn,by="IID")
phenCombo <- left_join(phenCombo,patcov,by="IID") 
phenEric <- phenCombo %>% dplyr::select(FID.x, IID, GM412, rnGM412, logGM412, QUARTILE ,meanCIPN8,ordCIPN8,biCIPN8,ord3CIPN8,exCIPN8,ageaudio.x,agediagnosis,CumlCispdose,CumlBLEOdose) %>% dplyr::rename(FID=FID.x,quartile=QUARTILE,ageaudio=ageaudio.x)
phentable <- phenEric %>% mutate(CumlBLEOdose=ifelse(is.nan(CumlBLEOdose),NA,CumlBLEOdose)) #should any CumlBLEOdose == 0? Check before Raynaud's GWAS
```

test run:  
`plink --bfile N88.forImputation --1 --pheno PtStudy.CIPN8.phenofile --pheno-name biCIPN8 --logistic --adjust --covar PtStudy.covfile --out PtStudy.preImpute.2015-11-24.biCIPN8ageaudio.Cispdose`

`plink --bfile N88.forImputation --1 --pheno PtStudy.CIPN8.phenofile --pheno-name exCIPN8 --logistic --adjust --covar PtStudy.covfile --out PtStudy.preImpute.2015-11-24.exCIPN8ageaudio.Cispdose`

`plink --bfile N88.forImputation --1 --pheno PtStudy.CIPN8.phenofile --pheno-name biCIPN8 --logistic --adjust --covar PtStudy.covfile --covar-name ageaudio --out PtStudy.preImpute.2015-11-24.biCIPN8ageaudio`

`plink --bfile N88.forImputation --1 --pheno PtStudy.CIPN8.phenofile --pheno-name exCIPN8 --logistic --adjust --covar PtStudy.covfile --covar-name ageaudio --out PtStudy.preImpute.2015-11-24.exCIPN8ageaudio`

####TO DO: code for ordinal regression phenotypes-->done, see 14_ordreg_GWAS.R

###Does ordCIPN8 associate with agediagnosis? CumlCispdose?
- only agediagnosis is signficant:
```{r cov}
phenofile <- ptcov.dir %&% "PtStudy.phenofile"
covfile <- ptcov.dir %&% "PtStudy.covfile"
pheno <- read.table(phenofile,header=T)
cov <- read.table(covfile,header=T) %>% dplyr::select(FID,starts_with("PC"))
pheno <- left_join(pheno,cov,by="FID")

#ordCIPN8 ~ agediagnosis
m<-polr(factor(pheno$ordCIPN8) ~ pheno$agediagnosis, Hess=TRUE)
p <- pnorm(abs(summary(m)$coefficients[, "t value"]), lower.tail = FALSE) * 2
or <- exp(summary(m)$coefficients[,"Value"])
ci <- exp(confint.default(m))
colnames(ci) <- c("lCI","uCI")
res <- cbind(summary(m)$coefficients[,],"OR"=or,"pval"=p)[1,]
res <- c(res, ci[1,])
res

#ordCIPN8 ~ CumlCispdose
m<-polr(factor(pheno$ordCIPN8) ~ pheno$CumlCispdose, Hess=TRUE)
p <- pnorm(abs(summary(m)$coefficients[, "t value"]), lower.tail = FALSE) * 2
or <- exp(summary(m)$coefficients[,"Value"])
ci <- exp(confint.default(m))
colnames(ci) <- c("lCI","uCI")
res <- cbind(summary(m)$coefficients[,],"OR"=or,"pval"=p)[1,]
res <- c(res, ci[1,])
res

#ordCIPN8 ~ agediagnosis + CumlCispdose
m<-polr(factor(pheno$ordCIPN8) ~ pheno$agediagnosis + pheno$CumlCispdose, Hess=TRUE)
p <- pnorm(abs(summary(m)$coefficients[, "t value"]), lower.tail = FALSE) * 2
or <- exp(summary(m)$coefficients[,"Value"])
ci <- exp(confint.default(m))
colnames(ci) <- c("lCI","uCI")
res <- cbind(summary(m)$coefficients[,],"OR"=or,"pval"=p)
res <- cbind(res[1:2,],ci)
res

#ordCIPN8 ~ agediagnosis + 10 PCs
covmat <- dplyr::select(pheno,agediagnosis,starts_with("PC"))
covmat <- as.matrix(covmat)
m<-polr(factor(pheno$ordCIPN8) ~ covmat, Hess=TRUE)
p <- pnorm(abs(summary(m)$coefficients[, "t value"]), lower.tail = FALSE) * 2
or <- exp(summary(m)$coefficients[,"Value"])
ci <- exp(confint.default(m))
colnames(ci) <- c("lCI","uCI")
res <- cbind(summary(m)$coefficients[,],"OR"=or,"pval"=p)
res <- cbind(res[1:dim(covmat)[2],], ci)
res

#ordCIPN8 ~ agediagnosis + CumlCispdose + 10 PCs
covmat <- dplyr::select(pheno,agediagnosis,CumlCispdose,starts_with("PC"))
covmat <- as.matrix(covmat)
m<-polr(factor(pheno$ordCIPN8) ~ covmat, Hess=TRUE)
p <- pnorm(abs(summary(m)$coefficients[, "t value"]), lower.tail = FALSE) * 2
or <- exp(summary(m)$coefficients[,"Value"])
ci <- exp(confint.default(m))
colnames(ci) <- c("lCI","uCI")
res <- cbind(summary(m)$coefficients[,],"OR"=or,"pval"=p)
res <- cbind(res[1:dim(covmat)[2],], ci)
res
```

###Pull SCIN
- Tinnitus: ordTin (C_Q27ERSP)
- Reduced Hearing: ordHL (C_Q27FRSP)
- Raynaud's: ordRay (mean C_Q27CRSP & C_Q27DRSP, take ceiling)
- also pull CumlBLEOdose
- add to PtStudy.phenofile
```{r}
fam <- read.table(my.dir %&% "genotypes/N88.forImputation.fam")
colnames(fam) <- c("FID","patno","father","mother","sex","pheno") 
fam <- fam %>% mutate(IID=as.character(patno))
demog <- read.sas7bdat(pheno.dir %&% "demographics.sas7bdat")
scin <- demog %>% dplyr::select(patno,C_Q27ARSP,C_Q27BRSP,C_Q27CRSP,C_Q27DRSP,C_Q27ERSP,C_Q27FRSP) %>% rename(Pain_Tingling_Feet=C_Q27ARSP,Pain_Tingling_Hands=C_Q27BRSP,White_Cold_Hands=C_Q27CRSP,White_Cold_Feet=C_Q27DRSP,ordTin=C_Q27ERSP,ordHL=C_Q27FRSP)
a<-scin
dim(a)

ray <- dplyr::select(scin,White_Cold_Hands,White_Cold_Feet)
ordRay <- ceiling(rowMeans(ray,na.rm=TRUE))
hist(ordRay)
table(ordRay)/sum(table(ordRay))
morepheno <- dplyr::select(scin,patno,ordTin,ordHL) %>% mutate(IID=as.character(patno),ordRay=ordRay)
hist(morepheno$ordTin)
table(morepheno$ordTin)/sum(table(morepheno$ordTin))
hist(morepheno$ordHL)
table(morepheno$ordHL)/sum(table(morepheno$ordHL))

phentable2<-left_join(phentable,morepheno,by="IID")
write.table(phentable2,file=my.dir %&% "genotypes/PtStudy.phenofile",quote=F,row.names = F,col.names = T)
```

